---
title: "final_R_reference"
author: "Katie Webster"
date: "2022-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages


To load packages you first need to install them

install.packages("palmerpenguins")
install.packages("here")


```{r}
#Then we can tell R we want to use these packages with either require() or library()

library(palmerpenguins)
library(here)

#renaming plamerpenguins

penguins <- data.frame(penguins)

summary(penguins)

````
If you have your own data that you would like to load into R, this is how you would do it
```{r}
##first we use here() which always goes to my base working directory (WD) of my current  R porject
here()
#now we use read.csv to read our .csv file with here to help direct the path. And name the data.frame we just uploaded using = or <-

ginkgo <-read.csv(here("data", "ginkgo_data_2021.csv"))

ginkgo
```
## Data Structure 

Create a vector of numbers:
```{r} 
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")

```

I can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec

## The print() function accomplishes the same task:
print(char_vec)

```
If I want to know how many characters are in a vector I use length()

```{r}
length(num_vec)
```
A matrix can be made using a vector

```{r}
matrix_num_vec<-matrix(data=num_vec,nrow=3,ncol=2)
# Specify number of rows, nrow, and number of columns, ncol
matrix_num_vec

#turning the mnatrix into a dataframe.
df_matrix<-data.frame(matrix_num_vec)

df_matrix
```

If I want to know how many rows in a dataset I use nrow(datasetname) and the funtion will return the number of rows. 
```{r}
nrow(ginkgo)
```
I can do the same for columns for ncol(datasetname)
```{r}
ncol(ginkgo)
```
To find out both the number of rows and columns I can use the function dim(datasetname)
```{r}
dim(ginkgo)
```


## Sub-setting 


creating a new subset of a data.frame
In this example we will use the the ginkgo data set and select only one column to be the new data.frame using $

```{r}
#data.frame$name of column you would like to subset

Gpetiole_length<-ginkgo$petiole_length

Gpetiole_length

```
Sometimes you may want to subset even further by selecting the row and or column we do this with [ ]'s

```{r}
# select first row of the ginkgo data by position

ginkgo[1,]

# select the element in row 2, column 3

ginkgo[[3,2]]

#select the 3rd column of the ginkgo data

ginkgo[3]

```
In this example we also use the subset() function to retrieve all the data for Adelie penguins (in the species column) from the peuguins dataset

```{r}
subset(penguins,species=="Adelie")
```



## Numerical Data Exploration

Sometimes we may want a quick summary of our data.frame 
for this we will use summary()
```{r}
summary(penguins)

# we get the summary of categorical and numerical values for each column 
```
Or we might want to take the mean on a specific column in our data.frame

```{r}
#here we calculate the mean of bill length in mm for all penguins in the penguin data set 
# We have to add na.rm = TRUE to exclude the missing data points from our computations 
# If we don't do this it will return NA to us
mean(penguins$body_mass_g,na.rm = TRUE)
```

We can easily calculate the standard deviation of a data set by using sd()
```{r}
#again we have to use na.rm = TRUE here to return results
sd(penguins$body_mass_g,na.rm = TRUE)

```


## Graphical Data Exploration 

LET'S GRAPH :) 

Scatter plot: Using the ginkgo data, create a scatter plot of max leaf depth (x) and max leaf width (y).

```{r}
# plot() is the generic function in r 
# plot(y-axis~x-axis , data=)
#col = color of the plot symbol
#pch = plotting symbol 
#cex = the number determines the size of text to symbols should be scaled 
#main = title 
#xlab = the name of the x-axis 
#ylab = the name of the y-axis 
#xlim = x-axis limits 
#ylim = y-axis limits 

plotleafxleaf<-plot(max_width~max_depth,data=ginkgo, main="Max leaf depth vs Max leaf width of Ginkgos",
   xlab="Max Leaf Depth (mm)", ylab="Max Leaf Width (mm)", pch=18, col="#339999", cex= 1.5, xlim=c(0,150), ylim=c(0,130) )

 
```
Making a Histogram 

We use hist() to create a histogram of penguin flipper lengths

breaks = controlling the number of bins, bars or cells in a histogram 

```{r}

hist(
  penguins$flipper_length_mm,
  main = "Penguin Flipper Length Frequency",
  col = "lightblue",
  border = "orange",
  xlab = "Flipper Length (in mm)",
  breaks= 8)

```
Making Boxplots 

We start with a basic boxplot 

```{r}
boxplot(ginkgo$petiole_length)
```
Conditional box plot with a continous variable of leaf width conditioned on if seeds were present or not in the ginkgo data set 

```{r}
boxplot(max_width ~ seeds_present, data = ginkgo)
```
Making a multipanel figure of 4 histograms We use 
par = creates multiple plots at once 
mfrow = define the multi plot set up with (row, column) number
```{r}
par(mfrow = c(2, 2))
hist(
ginkgo$max_width,
main = "Histogram of Ginkgo Leaf Max Width",
xlab = "Leaf Max Width (mm)")
hist(
ginkgo$max_depth,
main = "Histogram of Ginkgo Leaf Max Depth",
xlab = "Leaf Max Depth (mm)")
hist(
ginkgo$notch_depth,
main = "Histogram of Ginkgo Leaf Notch Depth",
xlab = "Leaf Notch Depth (mm)"
)
hist(ginkgo$petiole_length,
main = "Histogram of Ginkgo Petiole Length",
xlab = "Petiole Length (mm)"
)

```

## Distribution Functions

dnorm(x, mean, standare deviation) is a function that gives you the value of the probability density function of a normal distribution according to your variables of choice 

```{r}
dnorm(2,1.5,1)
```
pnorm(x,mean, standard deviation) is a function that gives you the value of the cumulative density function of the normal distribution according to your variables of choice 
x= vector of quantiles 
```{r}
pnorm(2,1.5,1)
```

qnorm(x, mean, standard deviation) is a function on that gives you the inverse result of the cumulative density function (pnorm)
x= the significance level to be used 

```{r}
qnorm(.95,1.5,1)
```
dbinom(x_bin, size, probability) is like dnorm but returns the value of the probability density function for a binomial distribution

```{r}
dbinom(2,10,.6)
```
pbinom(x_bin, size, probability) is like pnorm but returns value of the cumulative density function for a binomial distribution
```{r}
pbinom(2,10,.6)
```
qbinom(x_bin, size, probability) is like qnorm but returns value of the cumulative density function for a binomial distribution
```{r}
dbinom(2,10,.6)
```










